ARGoS-Blockchain interface

### 2020-09-11 article.tex
upd

\title{ARGoS-Blockchain interface}
\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}

@@

\begin{abstract}

  This technical report documents the installation and use of the
  ARGoS-Blockchain interface. This interface allows for conducting
  experiments in the robot swarm simulator ARGoS and the smart
  contract blockchain framework Ethereum. The blockchain part uses the
  container software Docker in order to facilitate the installation
  and the parallelization of the blockchain nodes. The use of
  blockchain-based smart contracts provides a framework for managing
  security issues in robot swarms. The interface is a convenient way
  to prototype blockchain-based control software in robot
  swarms. Byzantine robots, that is, robots that behave in an
  unintended way can disrupt the swarm behavior. In this technical
  report, we detail the motivation, setup, and usage of the
  ARGoS-blockchain interface.

\end{abstract}

@@

Several recent works have addressed the combination of blockchain
technology and swarm robotics
\cite{StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013}. This
combination proves useful for security, consensus finding, robot
economies, robot-as-a-service, and potentially many other applications
in robot swarms. However, conducting experiments involving robot
swarms and blockchain technology requires a suitable testbed. A widely
used framework consists of the Pi-puck robot platform running the
Ethereum blockchain network. However, not every researcher has access
to physical robots, obtaining and maintaining them be both
time-intense and cost-intense, not to mention the lengthy execution of
experiments. Therefore, a simulation testbed is needed to quickly try
and test different controllers and scenarios. The ARGoS robot
simulator \citep{PinTriOGr-etal2012:si} is the state-of-the-art
research platform to conduct simulations in swarm robotics. Such a
framework proves useful both for researchers who do not have access to
a suitable robot platform, such as the Pi-puck robot. Additionally, a
suitable platform is important before running time- and cost-consuming
experiments with real robots.

@@

The open-source software is available at GitHub: the blockchain module
is available at
\footnote{\url{https://github.com/Pold87/ARGoS-Blockchain-interface}}
and the ARGoS module is available at
\footnote{\url{https://github.com/Pold87/robot-swarms-need-blockchain}}.

@@

The following setup steps describe the installation of the
ARGoS-Blockchain interface using Linux as operating system.

@@

@@

\begin{verbatim}
$ sudo apt-get install cmake libfreeimage-dev libfreeimageplus-dev \
  qt5-default freeglut3-dev libxi-dev libxmu-dev liblua5.3-dev \
  lua5.3 doxygen graphviz graphviz-dev asciidoc
\end{verbatim}

2. Build and install ARGoS system-wide.
\begin{verbatim}
$ cd argos3/
$ mkdir build
$ cd build
$ cmake ../src
$ make
$ make doc
$ sudo make install
\end{verbatim}

3. Build and install the ARGoS-epuck plugin.

\begin{verbatim}
$ cd ../../argos3-epuck/
$ mkdir build
$ cd build
$ cmake ../src
$ make
$ sudo make install
\end{verbatim}

@@

\begin{enumerate}[leftmargin=*]
\item Create a base folder for storing the two modules of the
  ARGoS-Blockchain interface.
\begin{verbatim}
$ mkdir interface/
$ cd interface/
$ pwd
/home/vstrobel/software/interface
\end{verbatim}

\item Obtain the source code for the blockchain module.
\begin{verbatim}
$ git clone https://github.com/Pold87/ARGoS-Blockchain-interface
\end{verbatim}

\item Obtain the source code for the ARGoS module.
\begin{verbatim}
$ git clone https://github.com/Pold87/robot-swarms-need-blockchain
\end{verbatim}
\end{enumerate}

@@

\begin{enumerate}[leftmargin=*]
\item By default, the execution of Docker commands requires root user
  privileges. To facilitate the execution of Docker commands, it is
  possible to run Docker as non-root user (i.e., without
  \texttt{sudo}) as follows (see
  \url{https://docs.docker.com/engine/install/linux-postinstall/} for
  more detailed instructions).
\begin{verbatim}
$ sudo groupadd docker
$ sudo usermod -aG docker $USER
\end{verbatim}
 % New item
\item Obtain the Ethereum source code and store it in the folder
\verb|myowngeth/|. Please note that Ethereum is compiled locally
here. While there are Docker images for Ethereum, this allows for more
customization (e.g., if you want to change the size of the DAG file,
as explained in Section~\ref{sec:limited-hardware}).

\begin{verbatim}
$ cd ARGoS-Blockchain-interface/ && cd geth/
$ git clone https://github.com/ethereum/go-ethereum.git myowngeth/
\end{verbatim}
% New item
\item Create the Ethereum image and initialize the Docker swarm.

\begin{verbatim}
$ docker build -t mygeth .
$ docker swarm init
\end{verbatim}
% New item
\item Set the ARGoS folder in the file \verb|global_config_blockchain.sh|.

\begin{verbatim}
$ nano ../local_scripts/global_config_blockchain.sh
ARGOSFOLDER="/home/vstrobel/software/interface/robot-swarms-need-blockchain"
\end{verbatim}

\item In order to be able to mine, you need to create the DAG
  datasets. Creating the files requires approximately 2~GB disk space
  and the execution of the script can take take several minutes.

@@

\begin{enumerate}
\item Compile the ARGoS code.
\begin{verbatim}
$ cd /home/vstrobel/software/interface/robot-swarms-need-blockchain/
$ mkdir build/
$ cd build/
$ cmake ..
$ make
\end{verbatim}

\item Set the variable \verb|DOCKERBASE| in the file
\verb|global_config.sh| to the full path where the Blockchain module
(Module 1) repository is located on your computer, for example:

\begin{verbatim}
$ cd ..
$ nano global_config.sh
DOCKERBASE='/home/vstrobel/Documents/docker-geth-network/'
\end{verbatim}
\end{enumerate}

@@

Usually, the network is created when a swarm robotics experiment is
started, using one of the starter scripts in the folder
\verb|robot-swarms-need-blockchain/starters/|. Start with a plain
experiment to see if the installation was succesful. The command needs
to be executed from the base folder of the ARGoS part of the
blockchain interface.

\begin{verbatim}
$ bash starters/1_Plain.sh 3
\end{verbatim}

@@

\section{Modifications}

@@

First, check that the Docker container for the bootstrap node and the
individual containers for the robots are running:

\begin{verbatim}

\end{verbatim}

% Mention --no-trunc and how to access logs

@@

\begin{itemize}
\item \texttt{unlockAccount}: Unlock the coinbase account.
\item \texttt{startMining}: Start the mining process using one thread.
\item \texttt{stopMining}: Stop the mining process.
\item \texttt{addPeer}: Add a peer based on its enode.
\item \texttt{removePeer}: Add a peer based on its enode.
\item \texttt{scInterface}: 
\end{itemize}

>>>>

\title{ARGoS-Blockchain interface}
%\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}

@@

\begin{abstract}
  This technical report documents the installation and use of the
  ARGoS-Blockchain interface. This interface allows for conducting
  experiments in the robot swarm simulator ARGoS and the smart
  contract blockchain framework Ethereum. The blockchain part uses the
  container software Docker in order to facilitate the installation
  and the parallelization of the blockchain nodes. The use of
  blockchain-based smart contracts provides a framework for managing
  security issues in robot swarms. The interface is a convenient way
  to prototype blockchain-based control software in robot
  swarms. Byzantine robots, that is, robots that behave in an
  unintended way can disrupt the swarm behavior. In this technical
  report, we detail the motivation, setup, and usage of the
  ARGoS-blockchain interface.
\end{abstract}

@@

Several recent works have addressed the combination of blockchain
technology and swarm robotics
\cite{StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013,Cas2016:arxiv,PacStrDor2020:techreport-001,PacStrDor2020:techreport-007}. This
combination proves useful for security, consensus finding, robot
economies, robot-as-a-service, and potentially many other applications
in robot swarms. Being a young research field, it misses standardized
research platforms. However, conducting experiments involving robot
swarms and blockchain technology requires a suitable testbed. A widely
used framework consists of the Pi-puck robot platform running the
Ethereum blockchain network. However, not every researcher has access
to physical robots, obtaining and maintaining them be both
time-intense and cost-intense, not to mention the lengthy execution of
experiments. Therefore, a simulation testbed is needed to quickly try
and test different controllers and scenarios. The ARGoS robot swarm
simulator \citep{PinTriOGr-etal2012:si} is the state-of-the-art
research platform to conduct simulations in swarm robotics. Such a
framework proves useful both for researchers who do not have access to
a suitable robot platform, such as the Pi-puck robot. Additionally, a
suitable platform is important before running time- and cost-consuming
experiments with real robots.

@@

The open-source software is available at GitHub: the blockchain module
is available at
\footnote{\url{https://github.com/Pold87/AB-interface-Blockchain-module}}
and the ARGoS module is available at
\footnote{\url{https://github.com/Pold87/AB-interface-ARGoS-module}}.

@@

The following setup steps describe the installation of the
ARGoS-Blockchain interface using Linux as operating system. There is a
video tutorial at \url{https://www.youtube.com/watch?v=IqiTJK_uzUY}
that described the entire installation process, starting from a blank
Ubuntu 20.04 LTS installation.

@@

\subsection{Installation of Docker}

The installation of the Docker Engine on Ubuntu is described at
\url{https://docs.docker.com/engine/install/ubuntu/}. Below, you will
find information about how to control the Docker Engine as a non-root
user.

\begin{enumerate}[leftmargin=*]
\item By default, the execution of Docker commands requires root user
  privileges. To facilitate the execution of Docker commands, it is
  possible to run Docker as non-root user (i.e., without
  \texttt{sudo}) as follows (see
  \url{https://docs.docker.com/engine/install/linux-postinstall/} for
  more detailed instructions).
\begin{verbatim}
$ sudo groupadd docker
$ sudo usermod -aG docker $USER
$ newgrp docker
\end{verbatim}

\item Additionally, restarting the Docker engine via \verb|systemctl|
  requires sudo privileges. To control Docker as a normal user, add an
  entry in the sudoers file.
\begin{verbatim}
$ sudo visudo -f /etc/sudoers.d/docker
volker ALL= NOPASSWD: /bin/systemctl restart docker.service
volker ALL= NOPASSWD: /bin/systemctl start docker.service
volker ALL= NOPASSWD: /bin/systemctl stop docker.service
\end{verbatim}
\end{enumerate}

@@

\begin{verbatim}
$ sudo apt-get install cmake g++ libfreeimage-dev libfreeimageplus-dev \
  qt5-default freeglut3-dev libxi-dev libxmu-dev liblua5.3-dev \
  lua5.3 doxygen graphviz graphviz-dev asciidoc
\end{verbatim}

2. Build and install ARGoS system-wide.
\begin{verbatim}
$ cd argos3/
$ mkdir build
$ cd build
$ cmake ../src
$ make
$ make doc
$ sudo make install
$ sudo ldconfig
\end{verbatim}

3. Verify that the installation was succesful.

\begin{verbatim}
$ argos3 --version
ARGOS_VERSION=3.0.0-beta56
ARGOS_INSTALL_PREFIX=/usr/local
ARGOS_USE_DOUBLE=ON
ARGOS_WITH_LUA=ON
ARGOS_BUILD_FLAGS= -Wall -g -ggdb3
\end{verbatim}

4. Build and install the ARGoS-epuck plugin.

\begin{verbatim}
$ cd ../../argos3-epuck/
$ mkdir build
$ cd build
$ cmake ../src
$ make
$ sudo make install
$ sudo ldconfig
\end{verbatim}

@@

\begin{enumerate}[leftmargin=*]
\item Create a base folder for storing the two modules of the
  ARGoS-Blockchain interface.
\begin{verbatim}
$ mkdir AB-interface/
$ cd AB-interface/
$ pwd
/home/volker/software/AB-interface
\end{verbatim}

\item Obtain the source code for the blockchain module.
\begin{verbatim}
$ git clone https://github.com/Pold87/AB-interface-Blockchain-module.git \ 
  Blockchain-module
\end{verbatim}

\item Obtain the source code for the ARGoS module.
\begin{verbatim}
$ git clone https://github.com/Pold87/AB-interface-ARGoS-module.git ARGoS-module
\end{verbatim}
\end{enumerate}

@@

\begin{enumerate}
 % New item
\item Obtain the Ethereum source code and store it in the folder
\verb|custom-geth/|. Please note that Ethereum is compiled locally
here. While there are Docker images for Ethereum, this allows for more
customization (e.g., if you want to change the size of the DAG file,
as explained in Section~\ref{sec:limited-hardware}).

\begin{verbatim}
$ cd ARGoS-Blockchain-interface/ && cd geth/
$ git clone https://github.com/ethereum/go-ethereum.git custom-geth/
\end{verbatim}
% New item
\item Create the Ethereum image and initialize the Docker swarm.

\begin{verbatim}
$ docker build -t mygeth .
$ docker swarm init
\end{verbatim}
% New item
\item Next, we need to link the blockchain module and the ARGoS module. First, set the ARGoS folder in the file \verb|global_config_blockchain.sh|.

\begin{verbatim}
$ nano ../local_scripts/global_config_blockchain.sh
ARGOSFOLDER="/home/volker/software/AB-interface/ARGoS-module"
\end{verbatim}

\item Then, set the variable \verb|DOCKERBASE| in the file
  \verb|global_config.sh| of the ARGoS module (Module~2) to the full
  path where the blockchain module (Module~1) repository is located on
  your computer, for example:

\begin{verbatim}
$ nano /home/volker/software/AB-interface/ARGoS-module/global_config.sh
DOCKERBASE='/home/volker/software/AB-interface/Blockchain-module/'
\end{verbatim}

\item In order to be able to mine, you need to create the DAG
  datasets. Creating the files requires approximately 2~GB disk space
  and the execution of the script can take take several minutes. If
  you want to run the software on limited hardware, you can decrease
  the size of the DAG dataset by carrying out the modifcations
  described in Section~\ref{sec:limited-hardware}.

@@

\begin{enumerate}
\item Compile the ARGoS code.
\begin{verbatim}
$ cd /home/volker/software/AB-interface/ARGoS-module/
$ mkdir build/
$ cd build/
$ cmake ..
$ make
\end{verbatim}
\end{enumerate}

@@

Usually, the network is created when a swarm robotics experiment is
started, using one of the starter scripts in the folder
\verb|ARGoS-module/starters/|. Start with a plain
experiment to see if the installation was succesful. The command needs
to be executed from the base folder of the ARGoS part of the
blockchain interface.

\begin{verbatim}
$ bash starters/1_Plain.sh
\end{verbatim}

@@

\section{Modifications}
\label{sec:modifications}

@@

First, check that the Docker container for the bootstrap node and the
individual containers for the robots are running:
%
\begin{verbatim}
$ docker service ps ethereum_bootstrap --no-trunc
$ docker service ps ethereum_eth --no-trunc
\end{verbatim}
%

% \begin{verbatim}
% ID                          NAME                   IMAGE               NODE                DESIRED STATE       CURRENT STATE           ERROR               PORTS
% ik6isqahrmsgzkzkj4pbb2zd5   ethereum_bootstrap.1   mygeth:latest       volker-TERRA-PC     Running             Running 4 minutes ago   
% \end{verbatim}

If the value of \verb|DESIRED STATE| is \verb|RUNNING|, take a look at
the log files of the containers.

\begin{verbatim}
$ docker logs ethereum_bootstrap.1.ik6isqahrmsgzkzkj4pbb2zd5
\end{verbatim}

@@

\begin{itemize}
\item \texttt{unlockAccount}: Unlock the coinbase account.
\item \texttt{startMining}: Start the mining process using one thread.
\item \texttt{stopMining}: Stop the mining process.
\item \texttt{addPeer}: Add a peer based on its enode.
\item \texttt{removePeer}: Add a peer based on its enode.
\item \texttt{scInterface}: A generic smart contract interface.
\item \texttt{scReturn0(string function, long long wei)}: 
\end{itemize}

\subsection{Compilation}
\label{sec:compilation}

After changes to the ARGoS files are made, the code needs to be
recompiled.

\begin{verbatim}
$ cd build/
$ make
\end{verbatim}
